---
title: "Synthea Federated Cox PH Results"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
params:
  result: "latest"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

required <- c("jsonlite", "ggplot2")
missing <- required[!vapply(required, requireNamespace, logical(1), quietly = TRUE)]
if (length(missing)) {
  stop(
    "Install the required R packages first: install.packages(c(",
    paste(sprintf("'%s'", missing), collapse = ", "), "))"
  )
}
has_plotly <- requireNamespace("plotly", quietly = TRUE)
has_dt <- requireNamespace("DT", quietly = TRUE)

rmd_path <- knitr::current_input(dir = TRUE)
if (nzchar(rmd_path)) {
  demo_dir <- dirname(normalizePath(rmd_path))
} else if (basename(getwd()) == "synthea_nvflare") {
  demo_dir <- normalizePath(getwd())
} else {
  demo_dir <- normalizePath(file.path(getwd(), "synthea_nvflare"))
}
results_dir <- file.path(demo_dir, "results")
if (identical(params$result, "latest")) {
  latest <- jsonlite::fromJSON(file.path(results_dir, "latest_run.json"))
  result_dir <- file.path(demo_dir, latest$result_dir)
} else {
  result_dir <- file.path(results_dir, params$result)
}
manifest <- jsonlite::fromJSON(file.path(result_dir, "run_manifest.json"))
```

This report uses the best federated model and held-out test patients from each client.
The default data contain **simulated five-year mortality outcomes for demonstration**, so
the estimates must not be interpreted clinically.

## Run summary

```{r run-summary}
knitr::kable(data.frame(
  field = c("Run", "Cohorts", "Outcome", "Horizon (days)", "Held-out predictions", "Rounds"),
  value = c(
    manifest$run_name,
    paste(manifest$cohorts, collapse = ", "),
    manifest$outcome,
    manifest$horizon_days,
    manifest$patients_in_test_predictions,
    manifest$federated_rounds_reported
  )
))
```

```{r load-data, include=FALSE}
predictions <- read.csv(file.path(result_dir, "predictions.csv"), check.names = FALSE)
coefficients <- read.csv(file.path(result_dir, "coefficients.csv"), check.names = FALSE)
metrics <- read.csv(file.path(result_dir, "round_metrics.csv"), check.names = FALSE)

predictions$outcome <- ifelse(
  predictions$event == 1,
  "Event before horizon",
  "Censored/alive at horizon"
)
predictions$patient_short <- substr(predictions$patient_id, 1, 8)
predictions <- predictions[order(predictions$predicted_survival), ]
predictions$prediction_rank <- seq_len(nrow(predictions))
coefficients$direction <- ifelse(
  coefficients$coefficient > 0,
  "Higher hazard",
  "Lower hazard"
)
```

## Federated validation by round

```{r metrics-plot}
shown_metrics <- metrics[metrics$metric %in% c("c_index", "brier_horizon"), ]
shown_metrics$metric <- factor(
  shown_metrics$metric,
  levels = c("c_index", "brier_horizon"),
  labels = c("C-index", "Brier score")
)

metric_plot <- ggplot2::ggplot(
  shown_metrics,
  ggplot2::aes(round, value, color = metric, text = paste0(
    "Round: ", round, "<br>Metric: ", metric, "<br>Value: ", round(value, 3)
  ))
) +
  ggplot2::geom_line(linewidth = 0.8) +
  ggplot2::geom_point(size = 1.8) +
  ggplot2::geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey50") +
  ggplot2::coord_cartesian(ylim = c(0, 1)) +
  ggplot2::labs(x = "Federated round", y = "Metric", color = NULL) +
  ggplot2::theme_minimal()

if (has_plotly) {
  plotly::ggplotly(metric_plot, tooltip = "text")
} else {
  metric_plot
}
```

## Held-out survival predictions

```{r predictions-plot}
horizon_years <- unique(predictions$horizon_days) / 365
prediction_plot <- ggplot2::ggplot(
  predictions,
  ggplot2::aes(
    prediction_rank,
    predicted_survival,
    color = outcome,
    shape = client,
    text = paste0(
      "Patient: ", patient_short,
      "<br>Client: ", client,
      "<br>Predicted survival: ", round(predicted_survival, 3),
      "<br>Follow-up days: ", duration_days,
      "<br>Outcome: ", outcome
    )
  )
) +
  ggplot2::geom_point(size = 3) +
  ggplot2::coord_cartesian(ylim = c(0, 1)) +
  ggplot2::labs(
    title = paste0("Held-out ", horizon_years, "-year predictions"),
    x = "Patients sorted by predicted survival",
    y = "Predicted survival probability",
    color = NULL,
    shape = "Client"
  ) +
  ggplot2::theme_minimal()

if (has_plotly) {
  plotly::ggplotly(prediction_plot, tooltip = "text")
} else {
  prediction_plot
}
```

## Global Cox coefficients

Positive coefficients indicate higher modeled hazard per one-standard-deviation feature
increase; negative coefficients indicate lower modeled hazard. Missingness indicators
compare missing versus observed measurements. These are demo-model associations, not
causal or clinical effects.

```{r coefficient-plot}
coefficients$feature <- factor(
  coefficients$feature,
  levels = coefficients$feature[order(coefficients$coefficient)]
)
coefficient_plot <- ggplot2::ggplot(
  coefficients,
  ggplot2::aes(
    coefficient,
    feature,
    fill = direction,
    text = paste0(
      "Feature: ", feature,
      "<br>Coefficient: ", round(coefficient, 3),
      "<br>Hazard ratio: ", round(hazard_ratio, 3)
    )
  )
) +
  ggplot2::geom_col() +
  ggplot2::geom_vline(xintercept = 0, linewidth = 0.5) +
  ggplot2::labs(x = "Log-hazard coefficient", y = NULL, fill = NULL) +
  ggplot2::theme_minimal()

if (has_plotly) {
  plotly::ggplotly(coefficient_plot, tooltip = "text")
} else {
  coefficient_plot
}
```

## Patient-level results

```{r prediction-table}
shown_predictions <- predictions[c(
  "client", "patient_id", "duration_days", "event",
  "predicted_survival", "log_risk", "outcome_source"
)]
if (has_dt) {
  DT::datatable(
    shown_predictions,
    filter = "top",
    rownames = FALSE,
    options = list(pageLength = 10, scrollX = TRUE)
  ) |>
    DT::formatRound(c("predicted_survival", "log_risk"), digits = 3)
} else {
  knitr::kable(shown_predictions, digits = 3)
}
```
